Long-term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-23 article/article.tex
Commit 2024-06-23 16:14

Figure\nobreakspace{}\ref{fig:P-energy} shows the time series of the
yearly number of CE cases (each with duration of one minute), the yearly
mean OIR and the yearly excess irradiation for the period 1993 -- 2023,
together with corresponding linear trends. All three quantities show
increasing trends, most pronounced for the frequency of occurrence
(\(45.6\pm 21.9\,\text{cases}/\text{year}\)) and the excess irradiation
(\(116.9\pm 67.8\,\text{kJ}/\text{year}\)), which are also statistically
significant. In contrast the trend of the yearly mean OIR is negligible
(\(0.1\pm 0.1\,\text{W}/\text{m}^2/\text{year}\)) and of no statistical
significance. The average OIR for the entire period is about
\(39.9\,\text{W}/\text{m}^2\) with standard deviation of
\(2.7\,\text{W}/\text{m}^2\). The interannual variability of the data
about the trend lines is quite large. Furthermore, it tends to increase
with time (at least for the quantities of panels b and c), suggesting a
significant variability in cloud patterns over the area, possibly
associated to changes in climate.

@@

!!! monthly weigth !!!

Another interesting aspect of the CE events is their distribution within
the year. Figure\nobreakspace{}\ref{fig:relative-month-occurrences}
shows the monthly box and whisker plot of the CE number of occurrence
normalized with the highest median value that occurs in June, depicting
a clear seasonal cycle. Although CE events are present throughout the
year, the most active period is during May and June. During winter
(December -- February), the number of CE cases is about \(25\,\%\) of
the maximum, while in the intermediate months the number of occurrences
gradually ramps between the maximum and minimum. This seasonality is a
combined effect of different factors, among them the types of clouds,
their frequency of occurrence the seasonally varying relative position
of the sun, as well as the local landscape characteristics that may
influence the formation of clouds. Unfortunately, lack of detailed data
on cloud formation, type and location is not allowing further analysis.
The interannual variability of the monthly CE events is quite high as
manifested by the large monthly extremes, especially in the summer.

@@

\caption{Seasonal variability of the number of CE events normalized to the maximum occurring in June, in the form of a box and whisker plot. The monthly values have been also normalized to the relative abundance of valid GHI observations. The box contains the data between the lower $25\,\%$ and the upper $75\,\%$ percentiles. The thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values and the solid circles are outliers.}\label{fig:relative-month-occurrences}
\end{figure}

The distribution of the number of CE events as a function of OIR is
shown in Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently,
there is an inverse relation between the frequency of CE events and OIR
with an exponential-like decline. This is expected, as the stronger the
CE events become, the rarer are the conditions favoring the occurrence
of CE events. For the majority (over \(62\,\%\)) of the CE events the
OIR is below the longnterm average of \(39.9\,\text{W}/\text{m}^2\),
while about \(8.1\,\%\) of the events correspond to OIR larger than
\(100\,\text{W}/\text{m}^2\) and up to highest value of
\(412.4\,\text{W}/\text{m}^2\). This distribution is indicative of the
magnitude and the probability of the expected CE events over
Thessaloniki. Similar distribution of CE events, albeit with larger OIR
values, has been reported by \citet{Vamvakas2020} for the city of
Patras. This site is located \(2.5^\circ\) south of Thessaloniki and is
exposed to air masses coming mainly from the eastern Mediterranean
resulting in different cloud patterns, that may affect the
characteristics and magnitude of the CE events.

>>>>

Figure\nobreakspace{}\ref{fig:P-energy} shows the time series of the
yearly number of CE cases (each with duration of one minute), the yearly
mean OIR and the yearly excess irradiation for the period 1993 -- 2023,
together with corresponding linear trends. All three quantities show
increasing trends, most pronounced for the frequency of occurrence
(\(45.6\pm 21.9\,\text{cases}/\text{year}\)) and the excess irradiation
(\(116.9\pm 67.8\,\text{kJ}/\text{year}\)), which are also statistically
significant. In contrast the trend of the yearly mean OIR is negligible
(\(0.1\pm 0.1\,\text{W}/\text{m}^2/\text{year}\)) and of no statistical
significance. The average OIR for the entire period is
\(39.9\,\text{W}/\text{m}^2 \pm 2.7\,\text{W}/\text{m}^2\). The
interannual variability of the data about the trend lines is quite
large. Furthermore, the spread tends to increase with time (at least for
the quantities of panels b and c), suggesting a significant variability
in cloud patterns over the area, possibly associated to changes in
climate.

@@

Another interesting aspect of the CE events is their distribution within
the year. Figure\nobreakspace{}\ref{fig:relative-month-occurrences}
shows the monthly box and whisker plot of the CE number of occurrence
normalized with the highest median value, that occurs in June, depicting
a clear seasonal cycle. Although CE events are present throughout the
year, the most active months are May and June. During the winter
(December -- February), the number of CE cases is about \(25\,\%\) of
the maximum, while in the intermediate months, the number of occurrences
gradually ramps between the maximum and minimum. This seasonality is a
combined effect of different factors, among them the types of clouds,
their frequency of occurrence, the seasonally varying relative position
of the sun, as well as the local landscape characteristics that may
influence the formation of the clouds. Unfortunately, lack of detailed
data on cloud formation, type and location is not allowing further
analysis. The interannual variability of the monthly CE events is quite
high as manifested by the large monthly extremes, especially in the
summer.

@@

\caption{Seasonal variability of the number of CE events normalized to the maximum occurring in June, in the form of a box and whisker plot. The monthly values have been also normalized to the relative abundance of valid GHI observations. The box contains the data between the lower $25\,\%$ and the upper $75\,\%$ percentiles, the thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values and the solid circles are outliers.}\label{fig:relative-month-occurrences}
\end{figure}

The distribution of the number of CE events as a function of OIR is
shown in Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently,
there is an inverse relationship between the frequency of CE events and
OIR with an exponential-like decline. This is expected, as the stronger
the CE events are, the rarer the conditions favoring the occurrence of
CE events. For the majority (over \(62\,\%\)) of the CE events the OIR
is below the long-term average of \(39.9\,\text{W}/\text{m}^2\), while
about \(8.1\,\%\) of the events correspond to OIR larger than
\(100\,\text{W}/\text{m}^2\) and up to the highest value of
\(412.4\,\text{W}/\text{m}^2\). This distribution is indicative of the
magnitude and the probability of the expected CE events over
Thessaloniki. Similar distribution of CE events, albeit with larger OIR
values, has been reported by \citet{Vamvakas2020}, for the city of
Patras. This site is located \(2.5^\circ\) south of Thessaloniki and is
exposed to air masses coming mainly from the eastern Mediterranean
resulting in different cloud patterns, that may affect the
characteristics and magnitude of the CE events.
